---
title: "Introduction to tibbles"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tibble)
library(dplyr)
library(readr)
tutorial_options(exercise.timelimit = 60, exercise.cap = 'Sandbox')
knitr::opts_chunk$set(echo = TRUE)
mtcars_tbl <- rownames_to_column(mtcars)
```

## Introduction

<hr>

tibbles are a modern version of data frames that retain the good aspects (of data frames) while getting rid of the
frustrating and annoying parts.  In this module, we will learn how tibbles makes certain parts of the data analysis 
workflow easier by being different from data frames. 

### What will I learn?

<hr>

- what are tibbles?
- how are tibbles different from data frames?
- how to create tibbles?
- how to manipulate tibbles?


## Creating tibbles

<hr>

The first step in using tibbles is to learn how to create them. There are 
several ways of creating tibbles:

- use `tibble()` which is very similar to `data.frame()`
- convert another R object using `as_tibble()`
- use `tribble()` for manually entering the data

Let us start with `tibble()`. Creating tibbles using `tibble()` is similar to 
creating data frames using `data.frame()`. We need to supply the data and name 
for each column. Keep in mind that all the columns must be of the same length.

<!-- <br> -->

### Instructions

<hr>

Use the `tibble()` function to create a tibble with 2 columns. Follow the below 
instructions to create your first tibble:

- name the first column as `x` and assign to it the english alphabets using `letters`
- name the second column as `y` and assign to it the first 26 numbers
- Use `=` to assign values

```{r tibbles1, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Create tibble', exercise.lines = 5}
# create a tibble with 2 columns

```

```{r tibbles1-solution}
tibble(x = letters, y = 1:26)
```

## tibble Features

<hr>

Awesome! You just created a tibble. Before we continue learning the different 
ways to create tibbles, let us explore the key features of tibbles that 
differentiates it from data frames. 

A tibble will:

- never changes input's types
- never adjusts variable names
- never prints all rows
- never recycles vector of length greater than 1

## tibble Features 

<hr>

A tibble will never change the input's types. To understand this, let us create 
a `data.frame` and a `tibble` with the same underlying data. 

<!-- <br> -->

### Instructions

<hr>

- create a tibble with 2 columns
- name the first columns as `x` and assign it the english alphabets using `letters`
- name the second column as `y` and assign it the numbers `1:26`
- create a data frame with the same data and column names
- use the `=` operator to assign the values

```{r tibbles2, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Input Type', exercise.lines = 8}
# tibble


# data frame

```

```{r tibbles2-solution}
# tibble
tibble(x = letters, y = 1:26)

# data frame
data.frame(x = letters[1:10], y = 1:10)
```

You can observe that column `x` has been converted to `factor` in data frame 
while tibble does not change it from `character`. We will learn how to read in 
data as `character` or `factor` in the next module, where we learn how to 
import data into R.


## tibble Features 

<hr>

tibble will never adjust the names of variables. In the below example, we 
create a tibble and a data frame with the same variable name `order value`. 
In the case of data frame, it is modified to `order.value` where as tibble
retains the original name without modifying it. 

<!-- <br> -->

### Instructions

<hr>

- create a `data.frame` with column name `order value` and assign the value `10`
- create a tibble with column name `order value` and assign it the value `10`
- use `=` to assign the value to the column in the both the above cases
- while specifying the column names, enclose it with *`*

```{r tibbles3, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Variable Names', exercise.lines = 8}
# data frame


# tibble



```

```{r tibbles3-solution}
# data frame
data.frame(`order value` = 10)

# tibble
tibble(`order value` = 10)
```

## tibble Features 

<hr>

tibble will never print all the rows or columns unlike a data frame. It will 
print only 10 rows and only those columns that fit the output area. It will 
show the total number of rows and columns after printing the data. 

<!-- <br> -->

### Instructions

<hr>

- create a variable `x` and assign it the value 1 to 100 (`1:100`)
- create a variable `y` and assign it the first value of `letters` (`letters[1]`)
- create a tibble using `x` and `y`

```{r tibbles4, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Limit Rows', exercise.lines = 5}

```

```{r tibbles4-solution}
x <- 1:100
y <- letters[1]
tibble(x, y)
```

## tibble Features 

<hr>

tibble will never recycle any vector of length greater than 1 to avoid bugs or 
errors in the data.

<!-- <br> -->

### Instructions

<hr>

- create a tibble with 2 columns using `tibble()`
- name the first column as `x` and assign it the values `1:100`
- name the second column as `y` and assing it the english alphabets using `letters`
- run the code and read the error statement carefully
- now reassign the value `'a'` to the `y` column and run the code again


```{r tibbles5, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Recycle Vector', exercise.lines = 6}
# create a 2 column tibble

```

```{r tibbles5-solution}
x <- 1:100
y <- 'a'
tibble(x, y)
```

## Column Names

<hr>

Names of the columns in tibbles need not be valid R variable names. They can 
contain unusual characters like a space or a smiley but must be enclosed in 
ticks. 

<!-- <br> -->

### Instructions

<hr>

- create a tibble with 3 columns using `tibble()`
- the first column name must be space i.e. ` ` and assign it the value `'space'`
- the second column name must be `2` and assign it the value `'integer'`
- and the third column name must be `:)` and assign it the value `'smiley'`
- use the `=` operator to assign the values

```{r tibbles16, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Column Names', exercise.lines = 7}
# create a tibble with 3 columns 
tibble(
  ` ` = 'space',

)
```

```{r tibbles16-solution}
tibble(
  ` ` = 'space',
  `2` = 'integer',
  `:)` = 'smiley'
)
```



## Creating tibbles 

<hr>

Use `enframe()` to create tibbles from atomic vectors. If the elements of the 
vectors are named, `enframe()` will return a two column tibble i.e. one column 
for the element names and another for the values. In other cases, it will
return a one column tibble.

<!-- <br> -->

### Instructions

<hr>

- we have a created a vector of browser names
- pass it to `enframe()` to create a one column tibble
- pass `browser2` to `enframe()` and oberve the output

```{r tibbles11, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Vectors I', exercise.lines = 15}
# vector of browser names
browser1 <- c('chrome', 'safari', 'firefox', 'edge')

# create tibble from browsers


# named atomic vector
browser2 <- c(chrome = 40, firefox = 20, edge = 30, safari = 10)

# create tibble from browsers

```

```{r tibbles11-solution}
enframe(browser1)
enframe(browser2)
```

## tribble

<hr>

Another way to create tibbles is using `tribble()`: 

- it is short for transposed tibbles
- it is customized for data entry in code 
- column names start with `~` 
- and values are separated by commas

<!-- <br> -->

### Instructions

<hr>

We have provided the parital code to create a `tibble`. It included the column 
names and the separator for column names and data. You need to enter data for 
the first two rows. Follow the below instructions:

- enter `1`, `TRUE` and `'a'` for the first row
- enter `2`, `FALSE` and `'b'` for the second row
- enter a `,` after the first row

```{r tibbles13, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'tribble', exercise.lines = 8}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  
)
```

```{r tibbles13-solution}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  1, TRUE, 'a',
  2, FALSE, 'b'
)
```

## Membership Testing

<hr>

We can test if an object is a tibble using `is_tibble()`.

<!-- <br> -->

### Instructions

<hr>

- use `is_tibble()` to test if `mtcars` is a tibble
- coerce `mtcars` to a tibble using `as_tibble()` and then repeat the above step

```{r tibbles15, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'tribble', exercise.lines = 5}
is_tibble()
is_tibble()
```

```{r tibbles15-solution}
is_tibble(mtcars)
is_tibble(as_tibble(mtcars))
```

## Column Names

<hr>

Names of columns in tibbles need not be valid R variable names. They may 
contain unusual characters such as a smiley or a space but must be enclosed 
in ticks.

<!-- <br> -->

### Instructions

<hr>

- create a tibble with 3 columns using `tibble()`
- the column names must be 
    - space
    - 2
    - :)

```{r tibbles14, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'tribble', exercise.lines = 7}
tibble(
 


)
```

```{r tibbles14-solution}
tibble(
  ` ` = 'space',
  `2` = 'integer',
  `:)` = 'smiley'
)
```

## Add Rows

<hr>

Rows and columns can be added to a tibble using 

- `add_row()`
- `add_column()`

Let us say we have a tibble with data about website traffic driven by different browsers.

```{r browsers, echo=TRUE, collapse=FALSE}
browsers <- enframe(c(chrome = 40, firefox = 20))
browsers
```

Now, if we want to add another row of data for Internet Explorer, we can use `add_row()`

```{r browsers2, eval=TRUE, echo=TRUE}
add_row(browsers, name = 'IE', value = 30)
```

<!-- <br> -->

### Instructions

<hr>

- add data related to **Safari** browser to `browsers`
- set `name` to `'Safari'`
- and `value` to `10`

```{r tibbles17, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'tribble', exercise.lines = 5}

```

```{r tibbles17-solution}
browsers <- add_row(browsers, name = , value = )
browsers
```

## Add Rows

<hr>

In order to add the rows or columns at a specific index, use the following arguments

- `.before`
- `.after`

If we want to add the data at a particular row, we can specify the row number 
using the `.before` argument. Let us add the data related to Internet Explorer 
in the second row instead of the last row.

```{r browsers3, eval=TRUE, echo=TRUE}
add_row(browsers, name = "IE", value = 30, .before = 2)
```

If we want to add the data at a particular row, we can specify the row number 
using the `.after` argument. Let us add the data related to Internet Explorer 
after the first row.

```{r browsers4, eval=TRUE, echo=TRUE}
add_row(browsers, name = "IE", value = 30, .after = 1)
```

<!-- <br> -->

### Instructions

<hr>

- add data related to **Safari** browser to `browsers`
- set `name` to `'Safari'`
- and `value` to `10`
- add data after the second row

```{r tibbles18, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'tribble', exercise.lines = 5}

```

```{r tibbles18-solution}
add_row(browsers, name = "safari", value = 10, .after = 2)
```

## Add Columns

<hr>

`add_column()` adds a new column to tibbles.

<!-- <br> -->

### Syntax

<hr>

```{r tibble24a, eval=FALSE, echo=TRUE}
add_column(data, values)
```

<!-- <br> -->

### Instructions

<hr>

- add `visits` column to `browsers` data
- use `add_column()` to add the new column


```{r tibble24, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Add Columns', exercise.lines = 10}
# browsers
browsers <- enframe(c(chrome = 40, firefox = 20, edge = 30, safari = 10))

# visits
visits <- c(4000, 2000, 3000, 1000)

# add visits column to browsers

```

```{r tibble24-solution}
add_column(browsers, visits)
```

## Rownames

<hr>

The [tibble](https://tibble.tidyverse.org) package provides a set of functions 
to deal with rownames. Remember, `tibble` does not have `rownames` unlike 
`data.frame`. 

- use `has_rownames()` to check whether a data set has rownames
- use `remove_rownames()` to remove rownames from a data set
- use `rownames_to_column()` to convert the rownames to a column
- use `column_to_rownames()` to convert the first column in the data set to rownames

<!-- <br> -->

### Syntax

<hr>

```{r tibble25a, eval=FALSE, echo=TRUE}
has_rownames(data)
remove_rownames(data)
rownames_to_column(data)
column_to_rownames(data)
```

<!-- <br> -->

### Instructions

<hr>

- check if `mtcars` data has rownames using `has_rownames()`
- remove rownames from `mtcars` using `remove_rownames()`
- convert the rownames of `mtcars` into a column using `rownames_to_column()`
- we have created a tibble `mtcars_tbl`
- the rownames in `mtcars` is converted into a column in `mtcars_tbl`
- use `column_to_rownames()` to convert the first column of `mtcars_tbl` back to rownames


```{r tibble25, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Rownames', exercise.lines = 8}

```

```{r tibble25-solution}
has_rownames(mtcars)
remove_rownames(mtcars)
rownames_to_column(mtcars)
column_to_rownames(mtcars_tbl)
```

## Glimpse

<hr>

If you want to get a quick overview of the data set, use `glimpse()`. Another 
useful function is `has_names()`. It can be used to check if a tibble has a 
specific column.

<!-- <br> -->

### Syntax

<hr>

```{r tibble29a, eval=FALSE, echo=TRUE}
glimpse(data)
has_name(data, 'column_name')
```

<!-- <br> -->

### Instructions

<hr>

- use `glimpse()` to get an overview of `mtcars` data
- check if the column `cyl` is present in `mtcars` data set
- check if the column `gears` is present in the `mtcars` data set

```{r tibble29, exercise = TRUE, exercise.eval = FALSE, exercise.cap = 'Glimpse', exercise.lines = 6}

```

```{r tibble29-solution}
glimpse(mtcars)
has_name(mtcars, 'cyl')
has_name(mtcars, 'gears')
```


## Summary

### Manipulating tibbles

<hr>

- use `add_row()` to add a new row
- use `add_column()` to add a new column
- use `remove_rownames()` to remove rownames from data
- use `rownames_to_colum()` to coerce rowname to first column
- use `column_to_rownames()` to coerce first column to rownames

### Miscellaneous

<hr>

- use `is_tibble()` to test if an object is a tibble
- use `has_rownames()` to check whether a data set has rownames
- use `has_name()` to check if tibble has a specific column 
- use `glimpse()` to get an overview of data

### Creating tibbles

<hr>

- use `tibble()` to create tibbles
- use `as_tibble()` to coerce other objects to tibble
- use `enframe()` to coerce vector to tibble
- use `tribble()` to create tibble using data entry
